---
title: "Day03 – Assignment"
output:
  html_document:
    df_print: paged
---



# Part 1 --  Calculating p-value using a permutation test

This first part of the assignment is identical to the exercise we did in-class yesterday (Nov 16). I’m asking you to submit this R Notebook back to me with your comments/annotations on the code and the results, along with your interpretation and answers to a couple of questions at the end.


```{r}
### Don't need to run this commented code ###
# Generate some data

# effect_size <- 0.25
# stddev <- 0.5
# sample_size <- 20
#
# group1 <- round(rnorm(sample_size, mean = 0, sd = stddev)*10)
# group2 <- round(rnorm(sample_size, mean = effect_size, sd = stddev)*10)
#
# print(group1)
# print(group2)
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

# Data
group1 <- c(0, 5, 2, -2, 8,
            -6, 0, 0, -6, -3,
            -7, 4, -2, -2, 0,
            -4, -1, -8, 0, 6)

group2 <- c(-2, 2, -2, -5, 7,
            11, 6, 2, 1, 1,
            6, -2, -1, 7, 7,
            -2, 4, 3, -1, 17)

sample_size <- length(group1)
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

hist(group1,
     col = rgb(0, 0, 1, 0.25),
     breaks = seq(-20, 20, 2.5),
     xlim = c(-20, 20),
     main = "Histograms of group1 and group2")
hist(group2,
     col = rgb(1, 0, 0, 0.25),
     breaks = seq(-20, 20, 2.5),
     add = T)
abline(v = mean(group1))
abline(v = mean(group2))
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

signal <- mean(group2) - mean(group1)
noise <- sqrt((var(group2) / sample_size) + (var(group1) / sample_size))

test_statistic <- signal / noise
print(test_statistic)
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

num_permutations <- 10000
permuted_test_statistics <- {}
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

for(i in 1:num_permutations) {
  permuted_data <- sample(c(group1, group2))

  rand_group1 <- permuted_data[1:20]
  rand_group2 <- permuted_data[21:40]

  signal <- mean(rand_group2) - mean(rand_group1)
  noise <- sqrt((var(rand_group2) / 20) + (var(rand_group1) / 20))

  permuted_test_statistics <- c(permuted_test_statistics,
                                (signal / noise))
}
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

hist(permuted_test_statistics)
abline(v = test_statistic, col = "red", lwd = 2, lty = 2)
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

pvalue <- sum(permuted_test_statistics >= test_statistic) / num_permutations
cat("The pvalue based on a permutation test is", pvalue,"\n")
```

[ Add your notes here. ]

```{r}
t.test(group2, group1, alternative = "greater")
```


Question 1:
Define the p-value you just calculated in terms of this specific ‘humor medicine’ intervention experiment. The general definition is on slide 11 of the Day-03 lecture.

[ Write your answer here. ]


Question 2:
Is the p-value you calculated using the permutation test close to the p-value calcualted using the conventional t-test?

[ Write your answer here. ]



# Part 2 -- Effect of effect_size, variance, and sample_size on p-values

Like we discussed in class, apart from the null hypothesis (which need not always be equivalent to random chance), the p-value of a statistical test depends on:
1. Effect size,
2. Sample size, and
3. Variance within each group

To explore how these factors influence the p-value, I have written the code below to simulate data for two groups multiple times (just like the excercise we did in class), each time varying the `effect_size`, `std_deviation`, and `sample_size`, and calculating the p-value using a T-test.

Your tasks are the following:
1. Carefully examine and annotate the code by writing detailed comments at each step.
2. Examine the figure that is being produced and write a short paragraph about you observations on how these three quantities influence the p-value.


```{r}
library(tidyverse)
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

sample_data <- {}
es_sd_ss_pvalue <- {}

for(effect_size in seq(0.1, 1, 0.05)) {
  for(stddev in c(0.5, 1, 2)) {
    for(sample_size in c(5, 10, 20, 50, 100, 200, 500, 1000)) {
      group1 <- rnorm(sample_size, mean = 0, sd = stddev)
      group2 <- rnorm(sample_size, mean = effect_size, sd = stddev)

      sample_data <- rbind(sample_data,
                           cbind(rep(sample_size, sample_size),
                                 rep(effect_size, sample_size),
                                 rep(stddev, sample_size),
                                 group1,
                                 group2))

      ttest_out <- t.test(group1, group2)
      es_sd_ss_pvalue <- rbind(es_sd_ss_pvalue,
                            c(effect_size, stddev, sample_size, ttest_out$p.value))
    }
  }
}

colnames(es_sd_ss_pvalue) <- c("effect_size",
                               "std_deviation",
                               "sample_size",
                               "pvalue")
es_sd_ss_pvalue <- as_tibble(es_sd_ss_pvalue)
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

es_sd_ss_pvalue <- es_sd_ss_pvalue %>%
  mutate(pvalue_below_thresh = (pvalue < 0.05))
```

[ Add your notes here. ]

```{r}
# No need to add comments to this chunk that's just making the plot.
# However, it is useful to go through it to make sure you understand what is being plotted in the figure.

plot_es_sd_ss_pvalue <- es_sd_ss_pvalue %>%
  ggplot(aes(x = sample_size,
         y = effect_size,
         size = 2,
         #size = -log10(pvalue),
         color = pvalue_below_thresh,
         shape = pvalue_below_thresh)) +
  geom_point() +
  scale_shape_manual(values = c(4, 16)) +
  scale_color_manual(values = c("blue", "red")) +
  scale_x_log10() +
  facet_wrap(~std_deviation, nrow = 1) +
  theme_bw() +
  theme(legend.position = "none")

plot_es_sd_ss_pvalue
```

```{r}
# In case the figure above is squished, you can examine this PDF file
ggsave(filename = "effectsize_variance_samplesize_pvalue.pdf",
       plot_es_sd_ss_pvalue, height = 5, width = 12)
```


Question:
The figure contains 3 plots next to each other, one for each value `std_deviation`. In each plot, the `sample_size` values are along the x-axis and the `effect_size` values are along the y-axis. Each point is either a red dot or a blue cross depending on whether the p-value for the t-test between the two groups (for that combination of `effect_size`, `std_deviation` , and `sample_size`) is below 0.05 or not.

Examine the figure and write a short paragraph about you observations on how these three quantities influence the p-value.

[ Add your notes here. ]



# Part 3 -- P-hacking

P-hacking is the practice of collecting or selecting data or statistical analyses until nonsignificant results become significant.

You tasks:
1. Carefully examine and annotate the code by writing detailed comments at each step.
2. Describe your thoughts on what this coding exercise has to do with p-hacking.


[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

data <- rnorm(200)
hist(data)
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

attempts <- 0
pvalue <- 1
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

while(pvalue > 0.05) {
  attempts <- attempts + 1

  cases <- sample(data, 100)
  controls <- data[!data %in% cases]

  pvalue <- t.test(cases, controls)$p.value
}


cat("Congratulations! With p =", pvalue,
    "you achieved scientific success in", attempts, "attempts!\n") 
```

[ Add your notes here. ]


Question:
What this coding exercise has to do with p-hacking?

[ Write your answer here. ]
