---
title: "P-hacking"
output:
  html_document:
    df_print: paged
---

>Day03 – Assignment – Part 3

P-hacking is the practice of collecting or selecting data or statistical analyses until nonsignificant results become significant.

You tasks:
1. Carefully examine and annotate the code by writing detailed comments at each step.
2. Describe your thoughts on what this coding exercise has to do with p-hacking.


[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

data <- rnorm(200)
hist(data)
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

attempts <- 0
pvalue <- 1
```

[ Add your notes here. ]

```{r}
# Add comments next to each code chunk to describe the data analysis steps

while(pvalue > 0.05) {
  attempts <- attempts + 1

  cases <- sample(data, 100)
  controls <- data[!data %in% cases]

  pvalue <- t.test(cases, controls)$p.value
}


cat("Congratulations! With p =", pvalue,
    "you achieved scientific success in", attempts, "attempts!\n") 
```

[ Add your notes here. ]


Question:
What this coding exercise has to do with p-hacking?

[ Write your answer here. ]
